\documentclass{article}
\usepackage{array}
\usepackage{graphicx}
\usepackage{amsmath, amsthm, thmtools, mathdots, mathtools, amssymb}
\usepackage[pdfusetitle]{hyperref}
\usepackage[all]{hypcap} % Links hyperref to object top and not caption
\usepackage[nameinlink]{cleveref}   
\usepackage{makecell, multirow, booktabs}
\usepackage{caption, subcaption}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage[inline]{enumitem}
\usepackage{appendix}
\usepackage{biblatex}
\addbibresource{references.bib}
\hypersetup{ colorlinks, citecolor=black, filecolor=black, linkcolor=black, urlcolor=black, linktoc=all }
\newcommand*\rot{\rotatebox{90}}

\let\endtitlepage\relax

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lem}{Claim}


\begin{document}
    \begin{titlepage}
        \begin{center}
            {\LARGE Multiple Couriers Problem}
            \vspace*{1em}
            
            Valerio Costa, Luca Domeniconi, Claudia Maiolino, Tian Cheng Xia

            \centerline{\{valerio.costa, luca.domeniconi5, claudia.maiolino, tiancheng.xia\}@studio.unibo.it}
        \end{center}
    \end{titlepage}

    \thispagestyle{plain}

    \section{Introduction} \label{sec:intro}
    The problem of this project is known in the literature as the capacitated vehicle routing problem and can be easily proven to be NP-hard. We tackle this problem by virtually splitting it into two sub-problems: first we look for an assignment of the items to the couriers and then search for the routes of each of them (i.e., by solving multiple travelling salesman problems). To solve the latter, we follow the approach presented in \cite{vrp} where the route of a courier is modelled through the variables $P_d \in [1, n+1]$ with $d \in [1, n+1]$ defined as follows:
    \begin{equation}
        \label{eq:path_def}
        P_{d_1} = \begin{cases}
            d_2 & \text{iff the location $d_2 \neq d_1$ is visited immediately after $d_1$}\\
            d_1 & \text{iff $d_1$ is not part of the route of the courier}
        \end{cases}
    \end{equation} 
    With proper assignment and subtour elimination constraints, $P_d$ allows to define a Hamiltonian cycle that passes through the items that the courier delivers and the solution can be extracted by traversing the cycle starting from the depot $n+1$.

    The lower-bound of the objective function is common to all models and is defined as the maximum path cost that involves a single package:
    \begin{equation}
        \max_{p \in [1, n]} \left\{ D_{n+1, p} + D_{p, n+1} \right\}
    \end{equation}

    As upper-bound, also common to all models, we observed that it does not provide any improvement to the results. Nevertheless, we defined it as:
    \begin{equation}
        \sum_{d_1 \in [1, n+1]} \max_{d_2 \in [1, n+1]} D_{d_1, d_2}
    \end{equation}

    Symmetry breaking constraints are also common to all models. By considering couriers with the same capacity, the following constraints can be used to avoid symmetries:
    \begin{itemize}
        \item By imposing an ordering on the amount of assigned packages:
            \begin{equation}
                \label{eq:cp_symm_amount}
                \forall c_1, c_2 \in [1, m]: (c_1 < c_2 \land l_{c_1} = l_{c_2}) \Rightarrow Q_{c_1} \leq Q_{c_2}
            \end{equation}
            where $Q_c$ is the amount of packages delivered by the courier $c$.
        \item By imposing an ordering on the indexes of the assigned packages:
            \begin{equation}
                \label{eq:cp_symm_packs}
                \forall c_1, c_2 \in [1, m]: (c_1 < c_2 \land l_{c_1} = l_{c_2}) \Rightarrow A_{c_1} <_\texttt{lex} A_{c_2}
            \end{equation}
            where $A_c$ is an ordered vector containing the packages delivered by the courier $c$.
    \end{itemize}

    As the triangle inequality holds, we also identified an implied constraint that consists of imposing that each courier delivers at least a package (a short proof is provided in \Cref{sec:impl_proof}):
    \begin{equation}
        \label{eq:impl_constr}
        \forall c \in [1, m]: Q_c \geq 1  
    \end{equation}
    where $Q_c$ is the amount of packages delivered by the courier $c$.
    This obviously is applicable only if each courier is able to carry at least an item.

    All experiments were done using the same random seed and were run as workflows on GitHub Actions which provides two cores at 2.45 GHz and 7 GB of memory. To guarantee a safe margin for the Docker container to run, the actual usable memory was capped to 5 GB.

    The work has been completed in approximately one month and has been roughly split in the following way: Xia did the CP part, Costa worked on SAT, Domeniconi did the SMT models, and Maiolino completed the MIP part. The main difficulties we encountered are the following: 
    \begin{enumerate*}[label=(\roman*)]
        \item lack of proper documentation for many tools we used,
        \item difficulties to find a more efficient way to solve bigger instances,
        \item the need of time to run the experiments on all instances.
    \end{enumerate*}


    \input{./sections/cp.tex}
    \input{./sections/sat.tex}
    \input{./sections/smt.tex}
    \input{./sections/mip.tex}


    \section{Conclusions}

    We experimented several models using different methods and attempted to implement the same core idea across the whole project to make results comparable. Overall, all approaches are able to solve the smaller instances while, for bigger instances, only CP and SMT were able to at least produce a suboptimal solution by using proper search heuristics. Moreover, for this problem, we surprisingly noted that symmetry breaking constraints generally tend to, except in a few cases, worsen the results. To conclude, we can observe that, for this formulation of the problem, being able to guide the solver when exploring the search space is one of the most important factors to obtain good results.

    \printbibliography

    \begin{appendices}
        \section{Implied constraint proof} \label{sec:impl_proof}
        \begin{lem}
            Assuming that the capacity of each courier allows delivering at least a package, if there exists an optimal solution, then there exists an optimal solution where each courier delivers at least one package.
        \end{lem}
        \begin{proof}
            Let us assume that the optimal solution is $D_j$ and there is a courier, say $k_1$, which do not deliver any package. Let us also suppose that the courier $k_j$ is the one that covers the maximum distance $D_j$. If we assign one package that $k_j$ brings, say $i$, to $k_1$, then, due to the triangle inequality, the two new distances $D_1$, travelled by the courier $k_1$ with $i$, and $D_2$, travelled by the courier $k_j$ without $i$, are less or equal to $D_j$. In fact:
            \begin{equation}
                \begin{split}
                    D_1 = &D[\texttt{depot},i] + D[i,\texttt{depot}] \leq\\
                        & D[\texttt{depot}, i_1] + \dots + D[i_r, i] + D[i, i_s] + \dots + D[i_t, \texttt{depot}] = D_j.
                \end{split}
            \end{equation}
            \begin{equation}
                \begin{split}
                    D_2 = &D[\texttt{depot},i_1] + \dots + D[i_r,i_s] + \dots + D[i_t, \texttt{depot}] \leq\\
                    &D[\texttt{depot}, i_1] + \dots + D[i_r, i] + D[i, i_s] + \dots + D[i_t, \texttt{depot}] = D_j.
                \end{split}
            \end{equation}
            Therefore, there are the following cases:
            \begin{itemize}
                \item If $D_1 = D_2$, either both $k_1$ and $k_j$ cover the maximum distance or neither of them do, and another courier has a route of cost $D_j$.
                \item If $D_1 > D_2$, either $k_1$ covers the maximum distance or another courier that is not $k_1$ and $k_j$ does.
                \item If $D_1 < D_2$, same as above for $k_j$.
            \end{itemize}
            So, an optimal solution still exists and has as objective value $D_j$.
        \end{proof}
    \end{appendices}

\end{document}